---
title: "Lab 8"
output: html_notebook
---

```{r}
# load dplyr and rafalib since they are so useful
library(dplyr)
library(rafalib)
```

# Monte Carlo Simulations

## 1

```{r}
monte1 <- function(seed = FALSE){
  if(seed == TRUE){
    set.seed(1)
  }
  r_sample = rnorm(5)
  t_stat = sqrt(5) * mean(r_sample)/sd(r_sample)
  return(t_stat)
}
monte1.ans = monte1(TRUE)
print(monte1.ans)
```

##2
replicate 1000 times
```{r}
monte2 = function(){
  set.seed(1)
  B = 1000
  tstats = replicate(1000, monte1()) %>% c # must bind with c to be a vector
  length(tstats[tstats > 2])/length(tstats)
}

monte2.ans = monte2()
print(monte2.ans)
```

##3

```{r}
c(1-pt(2, df=4), monte2.ans)

monte3 <- function(){
  B = 100
  ps = seq(1/(B+1), 1-1/(B+1), len=B)
  qt(ps, df=4)
}

qqnorm(monte3())
qqline(monte3())
```

##5
```{r}
monte5gen <- function(N=15){
  B = 10000
  set.seed(1)
  tstats <- replicate(B,
  {X = sample(c(-1,1), N, replace=TRUE)
  sqrt(N)*mean(X)/sd(X)})
  
  ps = seq(1/(B+1), 1-1/(B+1), len=B)
  qqplot(qt(ps, N-1), tstats, xlim=range(tstats))
  abline(0,1)
  return(mean(tstats)-mean(ps))
}

monte5gen()

t.test(rnorm(15))$statistic
```

We can see that the tails of the QQplot ar far from the abline so thus the
answer is False

##6
```{r}
## since we developed a solution function that used N in #5 to look at N = 1000
## just run it with that parameter and look at the graph
monte5gen(1000)
````

The QQplot has tails that are in approximiate alignment with the abline, thus a
binary value generated distrubutation compared to a normally distrubuted one
show approximately equal t-statistics. Thus True

# Premutation Tests

Load our data

```{r}
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/babies.txt"
filename <- basename(url)
download.file(url, destfile=filename)
babies <- read.table("babies.txt", header=TRUE)
bwt.nonsmoke <- filter(babies, smoke==0) %>% select(bwt) %>% unlist 
bwt.smoke <- filter(babies, smoke==1) %>% select(bwt) %>% unlist
```

## 1
```{r}
permum1 <- function(){
  N = 10
  set.seed(1)
  ns <- sample(bwt.nonsmoke, N)
  s <- sample(bwt.smoke, N)
  obs <- mean(s) - mean(ns)
  dat <- c(s, ns)
  set.seed(1)
  perms <- function(){
    shuffle <- sample( dat )
    smokerstar <- shuffle[1:N]
    nonsmokerstar <- shuffle[(N+1):(2*N)]
    return(data.frame(smokerstar, nonsmokerstar))
  }
  s1 = c()
  s2 = c()
  rs = replicate(1000, {
    s = perms()
    s1 = rbind(s[1], s1)
    s2 = rbind(s[1], s1)
  })
  return(rs)
}

rs = permum1()


```
What is the premutation derived p-value for our observation?

