---
title: "Lab 2"
output: html_notebook
--- 

# Example 1
taking 4 samples. the first two samples are from a treatment group A and second two from a treatment group B.

thus X will look like

$$\mathbf{X} = \begin{bmatrix}
1 & 0 \\
1 & 0 \\
1 & 1 \\
1 & 1
\end{bmatrix}$$

and $\beta$ is:
$$\beta = [5, 2]$$

this is build around

$$\mathbf{Y} = \mathbf{X}\beta + \epsilon$$
thus we will be look to solve 

$$\mathbf{\beta} = (\mathbf{X}^\intercal\mathbf{X})^{-1} \mathbf{X}^\intercal\mathbf{Y}$$

which if find $\mathbf(X)\intercal\mathbf(Y)$ is:


```{r}
library(dplyr)

X <- matrix(c(1,1,1,1,0,0,1,1), nrow = 4)
betas <- matrix(c(5,2), nrow=2)

#set our rownames

rownames(X) <- c("a","a","b","b")
```
##1 

find the fitted value for A
```{r}
Axs <- X[which(rownames(X) %in% c("a")),]

#based on the code from pg 161 of the book "fitted" is:
# X %*% betahats
# thus the fitted value of A is Axs %*% betas
Axs %*% betas
```

##2

find the fitted value for B
```{r}
# and the fitted value for B is Bxs %*% betas
Bxs <- X[which(rownames(X) %in% c("b")),]
Bxs %*% betas
```

# Example 2

Now we have a trial with two treatements, B and C, and a control, A. Each group has two samples.

we will end up with B being:

$$\mathbf{B} = 
\begin{bmatrix}
1 & 1 & 0 \\
1 & 1 & 0
\end{bmatrix}$$

and our betas are:

$$\beta = \begin{bmatrix}
10\\
3\\
-3\\
\end{bmatrix}$$

##3

find the fitted value for b samples

```{r}

X2 = cbind(c(1), c(0,0,1,1,0,0), c(0,0,0,0,1,1))

row.names(X2) <- c("a","a","b","b","c","c")
betas2 <- c(10, 3, -3)

Bx2s <- X2[which(rownames(X2) %in% c("b")),]
Bx2s %*% betas2
```

##4

find the fitted value for c samples

```{r}

Cx2s <- X2[which(rownames(X2) %in% c("c")),]
Cx2s %*% betas2
```