---
title: "Lab 3"
output: html_notebook
---

This lab has me reviewing part of inference as from Stats and R ph525.1x. I am
going from using inference and applying matrix algebra work to it.

```{r}
## as always import dplyr first just in case I need it
library(dplyr)
```

# Example 1

Looking at least square estimates, let's do a simple physics problem!

```{r}

g = 9.8 # gravatational constant in meters/seconds

h0 = 56.67 # height in meters

v0 = 0 # intial velocity

n = 25 # number of measurements

tt = seq(0,3.4, len=n) # time in seconds as a vector

## calculate Y, rnorm introduces measurement error through creatng a normal
## distrubutation with n samples and sd of 1 and adding it to a polynomial
## equation with the height as the 0 degree term, velcoity in the 1 degree term
## and gravity constant and time squared in the 2 degree term (supposedly pret##
## common in phsyics *shrugs*)
set.seed(1)
y = h0 + v0 * tt - 0.5*g*tt^2+rnorm(n, sd=1)

## now let's take the LSE

## create matrix with 1 in one coliumn repeated, the tt in another and tt^2 in
## another representing  c b a of a ax^2+bx+c polynomia
iX = cbind(1, tt, tt^2)

## get crosspoduct of X (XtX) and solve then  matrix multiple transposed X
iA = solve(crossprod(iX)) %*% t(iX)
```

## 1

given defined A, whcich of the following is the LSE of g?

```{r}
-2*(iA %*% y)[3]
```

## 2

the estimate of g is a variable, use a monte carlo simulation (using replicate and sampling 10,000) to estimate the standard error of the sampling distrubutation of g with sd.

```{r}
rn = 100000
ll = function(){
  y = h0 + v0 * tt - 0.5*g*tt^2+rnorm(n, sd=1)
  -2*(iA %*% y)[3]
}
gs = replicate(rn, ll())
sd(gs)
hist(gs)
```

# Example 2

using father and son height data from UsingR, we are going to look at some of the correlation between them using LSE.

```{r}
library(UsingR)
fathers = father.son$fheight
sons = father.son$sheight

n = length(sons)
```

create Monte Carlo simulation (resampling disturubution with replacement):

```{r}

## create a function to be use by replicate to generate samples of father.son
## linear model coefficents/betas
gen_samps <- function(){
  # generate a list of  50 array-index values in a vactor based on the length of
  # the father height data
  index = sample(n,50)
  sampledat = father.son[index,]
  
  # sample father and sons as:
  # sfathers: sample of fathers
  # ssons: sample of sons
  sfathers = sampledat$fheight
  ssons = sampledat$sheight
  
  # get betahats that are coming lm (which is linear model) coefficents (coef)
  betas <- lm(ssons ~ sfathers)$coef
  return(betas)
}
set.seed(1)
samps <- replicate(10000, gen_samps())
```

##2

