---
title: 'Labwork 1: Pg. 13 #1-6, '
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Dplyr Excercises

## 1: 
load the data
```{r}
# uhis case driectly via the url from the github of ggplot2
url = "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/msleep_ggplot2.csv"

msleep =  read.csv(url)

#loiok at the head to see what the data looks liek
head(msleep)
```

# 2: 

filter only order Primates
```{r}
# load dplyr
library(dplyr)

primates <- filter(msleep, order == "Primates")

# show that primates data.frame only has primate in it
head(primates)

# get the summary statistics of  primates
summary(primates)

# finally how many records are there?
nrow(primates)
```

since knitr errored (filed a bug), here are some screenshots

![/home/crimson/develop/pubstatmeth/assignments/getting_started/Screenshot from 2018-01-22 09-28-02.png]
# 3:
What class is the primate subset

```{r}
# I expect it's a data.frame so let's check that theory with a equality
class(primates) == "data.frame"
```
Yep!

# 4:
select only total sleep, what class is it?

```{r}
# select sleep_total from msleep
totalSleepSub <- dplyr::select(msleep, sleep_total)

# since we know primate is a data.frame let's compare them
class(totalSleepSub) == class(primates)
```
Cool!

# 5:
select only totalsleep from primates as a numeric vector
```{r}
totalSleepVect <- select(msleep, sleep_total) %>% unlist

# True if totalSleepVect is a numeric vector
class(totalSleepVect) == "numeric"

# now the mean
mean(totalSleepVect)
```

#6:

use summarize to get the mean, removing the convert to vector step
```{r}
# use summarise based on the docs
select(msleep, sleep_total) %>% summarise(mean=mean(sleep_total))

```