---
title: "R Notebook"
output: html_notebook
---

Download Mice phenotypes
```{r}
library(downloader) 
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/mice_pheno.csv"
filename <- basename(url)
download(url, destfile=filename)
dat <- read.csv(filename) 

dat <- na.omit(dat)
```

```{r}
library(dplyr)

male_weights_chow <- dat %>% filter(Diet == "chow") %>% filter(Sex == "M") %>% select(Bodyweight) %>% unlist


head(male_weights_chow)
pop_x_bar <- mean(male_weights_chow)
```

#2

get the population sd via rafalib

```{r}
library(rafalib)
popsd(male_weights_chow)
```
#3

```{r}
set.seed(1)
sample_X_bar <- mean(sample(male_weights_chow, 25))
```

#4

make a vector of weights of male mice on the 'hf' diet

```{r}
male_weights_hf <- dat %>% filter(Sex == "M") %>% filter(Diet == "hf") %>% select(Bodyweight) %>% unlist
pop_y_bar = mean(male_weights_hf)
pop_y_bar
```

#5

get the population sd of bodyweights of male mice on the 'hf' diet

```{r}
popsd(male_weights_hf)
```

#6

take 25 samples and get the mean of the sample
```{r}
set.seed(1)
sample_Y_bar <- mean(sample(male_weights_hf, 25))
sample_Y_bar
```

#7

take the difference between $\bar{y_m}-\bar{x_m}$ and $\bar{Y_m} - \bar{Y_m}$

```{r}
set.seed(1)

abs(pop_y_bar - pop_x_bar)-abs(sample_Y_bar-sample_X_bar)
```

#8

Do the same as #7 for female mice

```{r}
# set the random seed to 1 to gaurantee results

# create two vectors:
# female_weights_hf -- the bodyweights of female mice on the "hf " diets
# femaile_weights_chow -- the bodyweights of female on the "chow" diets
female_weights_chow <- dat %>% filter(Sex == "F") %>% filter(Diet == "chow") %>% select(Bodyweight) %>% unlist
female_weights_hf <- dat %>% filter(Sex == "F") %>% filter(Diet == "hf") %>% select(Bodyweight) %>% unlist

#get the population means of both female_weights vectors
pop_x_bar_f <- mean(female_weights_chow)
pop_y_bar_f <- mean(female_weights_hf)

# take a sample of 25 from each vector and get the standard error of the mean
set.seed(1)
sample_X_bar_f <- sample(female_weights_chow, 25) %>% mean

set.seed(1)
sample_Y_bar_f <- sample(female_weights_hf, 25) %>% mean
```

Now, the difference from absolute value of $\bar{y_f} - \bar{x_f}$ and absolute value of $\bar{Y_f} - \bar{X_f}$
```{r}
abs(pop_y_bar_f - pop_x_bar_f) - abs(sample_Y_bar_f - sample_X_bar_f)

```

# CLT #2

#4

defining y as the body weights of male mice on the control diet ("chow") what proportion of the miace are within one $$\sigma$$ of the mean weight?

```{r}
male_y <- male_weights_chow

mean_male_y <- mean(male_y)
sd_male_y <- popsd(male_y)

proportion_n <- function(x, sd=sd_male_y, x_bar=mean_male_y){
  interval <- c(-sd, sd) + mean(x_bar)
  props = sapply(interval, function(y){mean(x <= y)})
  props[2] - props[1]
}

proportion_n(male_y)
```

#5

$2\sigma$ from the mean

```{r}
proportion_n(male_y, sd=(sd_male_y*2))
```

#6
$3\sigma$ from the mean

```{r}
proportion_n(male_y, sd=(sd_male_y*3))
```

#7

#8

```{r}
mypar(2,2)
y <- filter(dat, Sex=="M" & Diet=="chow") %>% select(Bodyweight) %>% unlist
z <- ( y - mean(y) ) / popsd(y)
qqnorm(z);abline(0,1)
y <- filter(dat, Sex=="F" & Diet=="chow") %>% select(Bodyweight) %>% unlist
z <- ( y - mean(y) ) / popsd(y)
qqnorm(z);abline(0,1)
y <- filter(dat, Sex=="M" & Diet=="hf") %>% select(Bodyweight) %>% unlist
z <- ( y - mean(y) ) / popsd(y)
qqnorm(z);abline(0,1)
y <- filter(dat, Sex=="F" & Diet=="hf") %>% select(Bodyweight) %>% unlist
z <- ( y - mean(y) ) / popsd(y)
qqnorm(z);abline(0,1)
```

```{r}
 y <- filter(dat, Sex=="M" & Diet=="chow") %>% select(Bodyweight) %>% unlist
set.seed(1)
avgs <- replicate(10000, mean( sample(y, 25)))
mypar(1,2)
hist(avgs)
qqnorm(avgs)
qqline(avgs)
mean(avgs)
sd(avgs)
```
  